Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2024-12-13 src/tex/ms.tex
[showyourwork] overleaf sync

\paragraph{Galactic chemical evolution models}
Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}. \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics. We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
% %
In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
\vspace{-1em}
% %\setlist[1]{itemsep=-4pt}
\begin{enumerate}
     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$. We treat these as star-independent and assume them to be constant across galactic environments and cosmic time. \footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G,2019A&A...626A.124M}.} We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).  
     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$. As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$. We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
\end{enumerate}

>>>>

\paragraph{Galactic chemical evolution models}
Our simulator is based on the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model \citep{Rybizki_2017}. \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics. We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.
% %
In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
\vspace{-1em}
% %\setlist[1]{itemsep=-4pt}
\begin{enumerate}
     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$. We treat these as star-independent and assume them to be constant across galactic environments and cosmic time. \footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{2014ApJ...796...75C,2016MNRAS.462.2832C,2019MNRAS.482..118G}}%,2019A&A...626A.124M}.} 
     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}).  
     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$. As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$. We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
\end{enumerate}
